---
title: "Chapter 5 Notes"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Start with the basic imports 

```{r imports}
library(Lahman)
library(tidyverse)
```

## 5.1 The Run Expectancy Matrix

An important concept in sabermetrics is the run expectancy matrix.

The matrix is made up of the average number of runs scored for a specific half-inning state. For example, we could have runners on 1st, 2nd, 3rd and combinations of them.  We also have0-3 outs. From there we can build 3 X 8 matrix. We don't include 3 outs becuse the half-inning is over and would always be 0.

Base runners  | 0 Outs        | 1 Out         | 2 Outs
------------- | ------------- | ------------- | -------------
100           |               |               |
110           |               |               |
101           |               |               |
111           |               |               |
000           |               |               |
001           |               |               |
011           |               |               |
010           |               |               |

 We denote the configuration of runners on base by the XXX values on the rows. The first slot is if a runner is on first, 2nd if a runner is on second, etc. Then we list the 8 possible combinations.

## 5.2 Runs Scored in the Remainder of the Inning

We will study the 2016 in this chapter and use the data supplied by the author.

read the column names:
 
 ```{r fields}
 fields <- read_csv("data/fields.csv")
 ```
 
Load the 2016 data and add the column names:

 ```{r data2016}
data2016 <- read_csv("data/all2016.csv",
                     col_names = pull(fields, Header),
                     na = character())
```
There's potential to score at each plate appearance with the chance being greater when runners are on base. 

